Analysis of observational data, to calculate the Delayed-Tau model parameters

### 2025-02-08 thesis.qmd
N-R+Markov-Chains+Monte_carlo

@@

@@

>>>>

The Newton-Raphson method is one of the most common numerical methods for estimating the root of a real-valued function. This method is based on the concept of linear approximation, where the function is locally approximated by its tangent line. It starts with initial guess of the root and iteratively refines the result using a formula that involves derivative of the function.

Let $x_n$ be the n-th approximation of the root and $f(x), f'(x)$ the function and the derivative, respectivelly. Then the refined approximation of the root $x_{n+1}$ is given by the formula

$$
x_{n+1} = x_n - \frac{f(x_n)}{f'(x_n)},\ n\in \mathbb{N}
$$

The iteration stops when $f(x_n)=0$, or after it reaches a certain number of iterations. If the initial estimation $x_0$ is close to the actual root, then the accuracy of the approximation increases rapidly with each iteration. If, however, the initial guess is poor, the model may fail to converge, which occurs when

$$
\left|\frac{f(x_n)f''(x_n)}{f'(x_n)^2}\right|>1.
$$

Therefore, while the Newton-Raphson method is a powerful tool in numerical analysis, due to its efficiency and ease of use, it is reliant on a good initial guess which is not always possible.

@@

Markov Chain Monte Carlo (MCMC) is a computer–driven sampling method which combines two properties: *Monte–Carlo* and *Markov chain*

A *Markov Chain* is a type of stochastic process that represents a sequence of events where the likelihood of each event depends solely on the state reached in the preceding event. In other words the future of each state depends only on the current state and not the past [@andrieuIntroductionMCMCMachine2003].

$$
\Pr(X_{n+1}=x\mid X_1=x_1, X_2=x_2, \ldots, X_n=x_n) = \Pr(X_{n+1}=x\mid X_n=x_n),
$$

if both conditional probabilities are well defined, that is, if $\Pr(X_1=x_1,\ldots,X_n=x_n)>0$.

A great example of such a chain is the board game snake and ladders. The dice has a uniform probabiliy  distribution across 6 stages (integers 1 to 6). The players have a position on the board, but their next position on the board is only based on the current position and the random roll of the dice. The one dimension random walk algorithm is also a Markov Chain.

The *Monte Carlo* method is a computational method that uses random values from a given probability distribution to approximate the features of the given distribution [@vanravenzwaaijSimpleIntroductionMarkov2018]. For example, if we want to calculate the mean of a normal distribution we can to draw a large number of random samples from the distribution, and calculate the sample mean of those, instead of calculating it directly.

@@

After the merging of the two catalogs we get a final sample of 3934 galaxies out of each only 288 are included in both catalogs and where used for the comparison. 1033 are unique galaxies in the UNGC and the remaining 2613 are unique in the HECATE catalogs. Additionally, 1842 galaxies have SFR data (778 from HECATE and 1137 from UNGC), and 3042 galaxies have Stellar Mass data (1605 from HECATE, with the rest estimated using a mass-to-light ratio of 0.82 as provided by HECATE). 1761 galaxies have both SFR and $M_*$ data.
